---
title: "Thermo Fisher raw file QC of BFabric Resource `r rawfileQC.parameter$resourceid`"
author: "Christian Trachsel"
date: "`r doc_date()`"
package: "`r pkg_ver('bfabricShiny')`"
abstract: >
  In this QC report a series of diagnostic plots are presented giving the reader a detailed overview of the raw file which was analyzed. This report should facitlitate the optimization of instrument methods or throubleshooting in case the results of a measurement are not as expecte. The report is focusing on MS parameters only. No direct chromatographic information is implemented as of today.
vignette: >
  %\VignetteIndexEntry{rawfileQC}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
output: 
  BiocStyle::pdf_document
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=8,
                      fig.height=8,
                      echo=FALSE,
                      warning=FALSE, 
                      message=TRUE)
```

\newpage

# Meta Data

```{r}
library(knitr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(bfabricShiny)

knitr::kable(rawfileQC.parameter$data.Info)
```

\newpage

# QC

## TIC versus RT
The Total Ion Chromatogram (TIC) as well as the Base Peak plot should present a quick overview about the signal distribution over the LC-MS run. It should help to descide if the proper amount of sample was loaded and give insights about the complexity of the sample.

```{r}
try(QC <- bfabricShiny:::.calc.master.scan(rawfileQC.parameter$data.QC))
try(QC <- bfabricShiny:::.calc.transient(QC))
```

```{r fig.cap='Total Ion Count (TIC) and Base peak plot.'}
try(bfabricShiny:::.TIC.BasePeak(QC))

```

\newpage

## Cycle Time and points over peak

The cycle time plot should help to descide if the current instrument method is suited for the LC gradient. A too long cycle time will result in to few points over the chromatographic peaks. This is especially problematic for a quantitative experiment. A estimation of the number of Ms data points over three fixed assumed peak widths (10 sec, 20 sec, 30 sec). 

```{r fig.cap='Cycle Time plot. The plot displays the calculated time between two consecutive Ms scans. The red horizontal line represents the 95ntil (95% of all datapoints are below the blue line). The blue curve displays a "gam" fit which indicates the cycle time trend over time.'}
try(bfabricShiny:::.cycle.time(QC))
```

\newpage

## distribution of selected m/z values

```{r fig.cap='Time resolved distribution of selecte precursor m/z values. The blue curve displays a "gam" fit which indicates the m/z trend over time.'}
try(bfabricShiny:::.mz.dist(QC))
```

\newpage

## Precursor mass frequency

```{r fig.cap='Frequency of peptide masses for each charge state. The plot displays frequency of all deconvoluted neutral mass values for all selected precursors according to their measured charge state.'}
try(bfabricShiny:::.mz.frequency(QC))
```

\newpage

## Frequency of precursor charge states

```{r fig.cap='Frequency of precursor charge states'}
try(bfabricShiny:::.charge.states(QC))
```

\newpage

## Cumulative precursor charges

```{r fig.cap='Cumulation of precursor charge percentages'}
try(bfabricShiny:::.charge.states.quantiles(QC))
```

\newpage

## Scan time plot

For HF instruments currently the scan to scan time difference is calculated. For Fusion data the "Elapsed Scan Time" value is 
plotted

```{r fig.cap='Scan to scan time difference plot. The red horizontal line indicates the theoretical transient lenght for the orbitrap resolution defined in the method. ITMS scans will not show this line.'}

try(bfabricShiny:::.scan.times(QC))

```

\newpage

## Injection time plot

```{r fig.cap='Injection time plot for Ms and Ms2 scans. Red horizontal lines displays the max. injection time value set in the instrument method. The Blue curves is a "gam" fit which indicated the injection time trend over time'}

try(bfabricShiny:::.injection.times(QC))
```

\newpage

## MS2 versus retention time

```{r fig.cap='Ms2 cycle load plotted agains retention time. The plot displays the density of Ms2 scans during the gradient.'}

try(bfabricShiny:::.ms2.vs.RT(QC))
```

\newpage

## MS2 cycle load

```{r fig.cap='Number of Ms2 scans between two consecutive Ms scans'}

try(bfabricShiny:::.ms2.frequency(QC))

```

\newpage

## Lock mass correction

```{r fig.cap='Lock mass correction vs RT. The plot displays the applied lock mass correction during the run. The two horizontal red lines indicate +/- 5ppm which indicates the FGCZ system recalibration limits.'}

try(bfabricShiny:::.lm.correction(QC))

```

\newpage

## MS datapoints over chromatographic peak

```{r fig.cap='Displays the number of Ms data points which are recorded with the current method over three theoretical chromatographic peak width. Plot is only a help. Actual peak width needs to be checked by XIC in the raw data.'}

try(bfabricShiny:::.ms.data.points(QC))
```

\newpage

# Session info

This report was compiled by using [bfabricShiny](https://github.com/cpanse/bfabricShiny) version `r packageVersion('bfabricShiny')` and `r R.version.string`.

# References

1. New RawFileReader from Thermo Fisher Scientific [planetorbitrap](http://planetorbitrap.com/rawfilereader#.WdOKzIqLnmG), 2017

2. https://github.com/cpanse/bfabricShiny
 
3. [bfabric resource `r rawfileQC.parameter$resourceid`](https://fgcz-bfabric.uzh.ch/bfabric/userlab/show-resource.html?resourceId=`r rawfileQC.parameter$resourceid`)

